---
title: "GoodPlot/BadPlot"
author: "Madigan Boborci"
format: 
  html:
    toc: true
    theme: cerulean
    mainfont: "Georgia"
    prefer-html: true
     
---
# Intro: 
This is my GoodPlot/BadPlot assignment. I am using the Serial Killer Dataset off Kaggle to attempt to make a beautiful visualization of the data, as well as a horrifying one. 

# Libraries 📚:
```{r}
#| warning: false
#| message: false


#load libraries
#loaded a bajillion libraries

library(tidyverse)
library(here)
library(ggmap)
library(tidytuesdayR)
library(emojifont)
library(patchwork)
library(devout)
library(devoutrgl)
library(triangular)
library(snowcrash)
library(cryogenic)
library(ggrgl)
library(maps)
library(plot3D)
library(RColorBrewer)
library(ragg)

```
# Data 🔢:

```{r}
#| warning: false
#| message: false

#Read in the data

highest_victim<-read.csv(here("GoodPlotBadPlot", "Data","SerialKiller", "Highest_victim_count.csv"))
victim30<-read.csv(here("GoodPlotBadPlot", "Data", "SerialKiller","15_to_30_victim_count.csv"))
victim14<-read.csv(here("GoodPlotBadPlot", "Data","SerialKiller", "5_to_14_victim_count.csv"))
victim5<-read.csv(here("GoodPlotBadPlot", "Data", "SerialKiller", "Lessthan_5_victim_count.csv"))


#take a gander at the data
#head(highest_victim)
#head(victim30)
#head(victim14)
#head(victim5)

victim14$Proven.victims<-as.integer(victim14$Proven.victims) #saw in head that one column was a character rather than an integer like the rest of the datasets


#can't use join because trying to merge vertically not horizontally
serial_killers <- rbind(highest_victim, victim30, victim14, victim5)
View(serial_killers)
```
Now that we have data to work with, we can make some plots.


# Good Plot! 😇
This plot is going to be an attempt to be a beautiful visualization of the data.

The first chunk will clean up some of the data
```{r} 
#| warning: false
#| message: false

#GOOD PLOT
#Now we have data, figure out what to do with it 
#need to split years active into start and end years, drop na's etc

killers_clean<-serial_killers%>%
  separate(
    col = Years.active,  # The column to split
    into = c("StartYear", "EndYear"), # Names of the new columns
    sep = " to ",          
    convert = TRUE)%>%
    drop_na()%>%
  write_csv(here("GoodPlotBadPlot", "Output", "killers_clean.csv"))

View(killers_clean)

```

Then we can prep that data specifically for the plot.
```{r}
Killers_countries <- killers_clean %>%
  
  filter(!grepl("\n", Country)) %>%   #drops all values that were listed as multiple countries
  group_by(Country) %>%
  mutate(Killer_Count = n())%>%     #make column for number of ind killers per country
distinct(Country, Killer_Count) %>% #makes mini dataset with just country and # killers
  slice_max(Killer_Count, n = 10, with_ties = FALSE)%>%     #Take top 10 highest count of killers 
arrange(desc(Killer_Count)) %>%             #arrange in descending order for plot
  head(10) 

View(Killers_countries)
```
 
Then this chunk creates the plot.
```{r}
#bar plot of amt of serial killers per country (top 10)

Killer_Plot <- Killers_countries %>%
  ggplot(aes(x = reorder(Country, Killer_Count), y = Killer_Count, fill=Killer_Count))+   #make countries go in order of no. of killers
  geom_col(color = "black",width=0.5) +
  scale_fill_gradientn(colors = brewer.pal(9, "Reds"))+  
  coord_flip() +    #I like it visualized with flipped axes
  labs(
    title = "Countries with the Most Serial Killers",
    x = "",
    y = "Number of Killers",
    subtitle="Serial Killers from 1900 to Present",
    fill="No. of Serial Killers") +
  theme_minimal() +
    theme(
    plot.title = element_text(hjust = 0.5, face = "bold"))
  

Killer_Plot
```

AND


This chunk preps the data we need to make the plot.
```{r}
#get data for all 10 countries with the most killers
top_10_countries<- killers_clean %>%
  filter(!grepl("\n", Country)) %>%   #remove values having more than 1 country listed
  count(Country) %>%                   
  slice_max(n, n = 10, with_ties = FALSE) %>%
  pull(Country) # Extracts a vector of the top 10 country names   

Killer_victims <- killers_clean %>%
  filter(Country %in% top_10_countries) %>%   #Uses the countries we selected above
  group_by(Country) %>%
  mutate(Killer_Count = n()) #makes a row for killer count but also keeps the rest of the info
     
```

This chunk creates the plot- the bar graph above becomes a visualization of the scale legend below.
```{r}
Killer_BoxPlot <- Killer_victims%>%
  ggplot(aes(x = reorder(Country, -Killer_Count), y = Proven.victims, fill= Killer_Count)) +
  geom_boxplot(color = "black",alpha=0.7) +
  scale_fill_gradientn(colors = brewer.pal(9, "Reds"))+   #choosing red for my legend scale bar
  labs(
    title ="Proven Victim Count by Country",
    subtitle="Countries with the most Serial Killers",
    x = "",
    y = "Number of victims",
    caption="Dataset from Serial Killers Data on Kaggle",
    fill="No. of Serial Killers") +
 
    theme(
    plot.title = element_text(hjust=.25, face = "bold"))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 10))+
  theme(axis.title=element_text(size=10))+
  theme(plot.title=element_text(size=14,
                                face="bold",
                                hjust=0))+
  theme(axis.text.x = element_text(angle=45,hjust = 1)) 


Killer_BoxPlot

```


This chunk stitches together the two plots I have already made, so they can be presented together. 
```{r fig.cap="GoodPlot", fig.width=10,fig.height=7}
#|message: false

#patched together? necessary?
(Killer_Plot/Killer_BoxPlot) +
  plot_annotation(tag_levels='A')  #putting labels on each subplot
ggsave(here("GoodPlotBadPlot","Output","GoodPlot_AB.png"),width=7,height=5)
```

This plot is considered a good plot because it follows best data practices as defined in _The American Statistician-Data Organization in Spreadsheets_ by Broman and Woo. These practices include **Good Naming Conventions**, **No Empty Cells**, and **One Item per Cell** and **Save Data as Plain Text Files**. I also created a **Data Dictionary**, to ensure people can understand what exactly the data are referring to. As a visualization, this plot follows the best practices we learned in class. Some of these aspects are

- Clean and visible plot and axis titles 
- No unnecessary colors   
- Clear legend if necessary  
- Decluttered  


  
# Bad Plot 🤢
Now I will make an atrocity, potentially more frightening than this dataset's contents.
I apologize for what you are about to see.

```{r fig.cap="BadPlot", message=FALSE, warning=FALSE}


#Now attempting a bad plot

#This is how to save- was put in after plot was created but needed to run first
#png(filename = "/Users/madiganboborci/Desktop/Repositories/Boborci/GoodPlotBadPlot/Output/BadPlot.png", 
   # width = 800, 
    #height = 600, 
   # res = 100) 

Killers_bad<-killers_clean %>%
  filter(!grepl("\n", Country)) %>%    #removing values with multiple countries
  group_by(Country) %>%              
  mutate(Killer_Count = n())%>%       #add column for serial killers per country
  mutate(StartYear = as.numeric(as.character(StartYear)))%>%    #needed all variables to be numbers
  mutate(EndYear=as.numeric(as.character(EndYear)))



with(Killers_bad, text3D(StartYear, Killer_Count, EndYear,        #define all axes
  labels = rownames(Killers_bad), colvar = Proven.victims,        #set legend
  col = heat.colors(100), theta = 60, phi = 20,                   #choose legend colors
  xlab = "Year Start", ylab = "Killer Count", zlab = "Year End",  #labels for axes
  main = "Serial Killers Worldwide", cex = 0.6,                   #Title
  bty = "u", ticktype = "detailed", d = 2,                        
  clab = c("Proven Victims"), adj = 0.5, font = 6,              #Label for legend
col.panel="blue",col.grid="black"))                             #commit atrocity

#dev.off() also for saving plot

```
There are many things that makes this plot a disgrace. First of all, it is **_3D_**. As we were taught in class, this is a general no-no for data visualization. It is unnecessarily complex and rarely displays data accurately. I also have two time metrics plotted on different axes, which is not intuitive at all. There are far too many variables beig analyzed here, with 3 axes and a scaled legend. My axees are not legible, and they are all titled but the titles aren't all legible either. Given the amount of metrics I have, the numbers within the graph are also corresponding to relationships between axes, no longer displaying the value those labels should represent. Lastly, my colors. There's just no reason for it to be bright blue. It doesn't contribute positively to the aestheitcs, it doesn't correlate to data, and it's just distracting. 











