---
title: "Week 7 Homework"
author: "Madigan Boborci"
date: "2025-10-7"
date-modified: today
format:
  html:
    toc: true
    theme: cerulean
    mainfont: "Georgia"
    prefer-html: true
    message: false 
    warning: false 
---
## Introduction
  This Quarto doc is to do the homework: making a map with some **Tiny Tuesday** data.
  
##### First I need to load my libraries
  
```{r}
#load libraries

#| warning: false
#| messgae: false
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(ggmap)
library(tidytuesdayR)
library(emojifont)
library(ggspatial)

```
##### Then I can read in the data I chose: **BIGFOOT 👣**
```{r}
#Read in data 
#| warning: false
#| messgae: false

tuesdata <- tidytuesdayR::tt_load('2022-09-13')
tuesdata <- tidytuesdayR::tt_load(2022, week = 37)

bigfoot <- tuesdata$bigfoot%>%
write_csv((here("Week_07", "Data", "bigfoot.csv"))   )
```


```{r}
#Look at data
View(bigfoot)
```

# Data Analysis
  I want to see where Bigfoot sightings have occurred by county in Illinois, my home state. 
  I need to  
-   make a map of Illinois 
-   change the data so the column county matches  
-   join the data  
-   overlay the sightings as a point layer? or heat map?  
-   make map pretty  
  
##### This chunk is for figuring out how to isolate and map Illinois
```{r}
#|message: false
#|warning: false

#Actually want to look at just Illinois
states<-map_data("state")
head(states)
county<-map_data("county")
head(county)

illinois<-county%>%
  filter(region=="illinois")
```
##### Now map it
```{r}
ggplot()+
  geom_polygon(data=illinois, aes(x=long,y=lat, group=group),fill="darkgrey",color="lightgrey")+
  guides(fill=FALSE)+
  coord_map()+
  theme_void()

```
##### Actually want to emphasize my own county
```{r}
kane<-county%>%
  filter(region=="illinois")%>%     #filter for Illinois and Kane County
  filter(subregion == "kane")

#map just the county to check
ggplot()+
  geom_polygon(data=kane, aes(x=long,y=lat, group=group))+
  guides(fill=FALSE)+
  coord_map()+
  theme_void()

```

##### Include highlighted county data on Illinois map
```{r}
ggplot()+
  geom_polygon(data=illinois, aes(x=long,y=lat,group=group,),fill = "linen", color="lightgrey")+
  geom_polygon(data=kane, aes(x=long,y=lat,group=group),fill="yellow")+
 coord_map()+
  theme_void()
```


##### Figure out how to fix up this bigfoot data
```{r}
#only care about state and county, lat and long

bigfoot_clean<-bigfoot%>%
  drop_na()%>%
  select(state,county,latitude,longitude)%>%   #select data wee need
  filter(state=="Illinois")%>%
  filter(longitude!="-86.13757")    #got rid of random typo longitude that made a floating point

```

##### Find data for Illinois
```{r}
head(county)
head(bigfoot_clean) #using these to double check the column names 

Bigfoot_illinois<-bigfoot_clean%>%
  rename(region=state, subregion=county)%>%  #change column names to match
  left_join(county, bigfoot_clean, by=c("region"))%>%
  write_csv(here("Week_07", "Output", "Bigfoot_illinois.csv"))   

#View(Bigfoot_illinois)

```





##### Now I can put the Bigfoot sightings data on top of the map I made

```{r}
ggplot()+
  geom_polygon(data=illinois, aes(x=long,y=lat,group=group,),fill = "linen", color="lightgrey")+
  geom_polygon(data=kane, aes(x=long,y=lat,group=group),fill="yellow")+
  geom_point(data=Bigfoot_illinois,
             aes(x=longitude, y=latitude, color=subregion.x))+    #colored points by county 
  guides(color="none")+
 coord_map()+
  theme_void()


```

##### I decided I wanted to include my hometown as well, so this chunk explores how to add another point to the map.  
```{r}
#Now an attempt to add a fun point where I grew up
geocode("Saint Charles, IL")
#long= -88.30869 and lat= 41.91419

StCharles<-data.frame(long= -88.30869,  lat=41.91419, name="home") #Make a point on my hometown
```


### All together now
```{r}
#Same code as above to see if adding this point works
ggplot()+
  geom_polygon(data=illinois, aes(x=long,y=lat,group=group,),fill = "linen", color="lightgrey")+
  geom_polygon(data=kane, aes(x=long,y=lat,group=group),fill="yellow")+
  geom_point(data=Bigfoot_illinois,
             aes(x=longitude, y=latitude, color=subregion.x))+
  guides(color="none")+
  geom_point(data=StCharles,aes(x=long, y=lat),shape="triangle", color="red",fill="red", size=3)+
 coord_map()+
  theme_void()

```

##### Create final map
```{r}
#| label: fig-bigfoot     #use tbl to label table
#| fig-cap: "Bigfoot sightings across counties in Illinois"
ggplot()+
  geom_polygon(data=illinois, aes(x=long,y=lat,group=group,),fill = "linen", color="lightgrey")+
  geom_polygon(data=kane, aes(x=long,y=lat,group=group),fill="yellow")+
  geom_point(data=Bigfoot_illinois,
             aes(x=longitude, y=latitude, color=subregion.x))+
  guides(color="none")+
  geom_point(data=StCharles,aes(x=long, y=lat),shape="triangle", color="red",fill="red", size=3)+
 coord_map()+
  theme_void()+
  scale_color_viridis_d()+
  annotation_scale(bar_cols=c("black","white"),
                   location="bl")+ #puts scale bar in bottom left
  annotation_north_arrow(location="tl")+ #puts direction arrow top left
  coord_sf(crs=4326) #GPS coordinate reference system- this is typical but depends on gps
ggsave(here("Week_07","Output","Bigfoot_IL.pdf"))

```
  
## Summary:
This map, @fig-bigfoot shows recorded Bigfoot Sightings in Illinois, and how close my hometown is to them. It includes a north arrow and a scale bar as well.  

  





